# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
activity_data <- read.csv(unz('activity.zip',filename='activity.csv'))
head(activity_data)
summary(activity_data)
```


## What is mean total number of steps taken per day?

```{r}
steps_by_day <- xtabs(steps ~ date, data=activity_data, na.action=na.omit)
head(steps_by_day)
hist(steps_by_day, )
```

Summary statistics:

* Mean of total steps per day: **`r mean(steps_by_day)`**
* Median of total steps per day: **`r median(steps_by_day)`**

```{r, echo=TRUE}
mean(steps_by_day)
median(steps_by_day)
```


## What is the average daily activity pattern?

```{r}
avg_steps_per_interval <- aggregate(steps ~ interval, data=activity_data, FUN=mean, na.action=na.omit)
head(avg_steps_per_interval)
plot(avg_steps_per_interval, type="l")
```

Interval **`r avg_steps_per_interval$interval[which.max(avg_steps_per_interval$steps)]
`** contains the maximum number of steps, on average across all the days in the dataset.

## Imputing missing values

Total numer is missing step values in the data set is: **`r sum(is.na(activity_data$steps))`**

Filling-in non avaliable step values by using the average value of for the interval across all days.

```{r}
merged_data <- merge(activity_data,avg_steps_per_interval, by='interval')
complete_activity_data <- data.frame(
    steps = ifelse(is.na(merged_data$steps.x), merged_data$steps.y, merged_data$steps.x),
    date = merged_data$date, 
    interval = merged_data$interval
)
head(completed_activity_data)
```


```{r}
complete_steps_by_day <- xtabs(steps ~ date, data=complete_activity_data, na.action=na.omit)
head(complete_steps_by_day)
hist(complete_steps_by_day)
```


Summary statistics:
* Mean of total steps per day: **`r mean(complete_steps_by_day)`**
* Median of total steps per day: **`r median(complete_steps_by_day)`**

```{r, echo=TRUE}
mean(complete_steps_by_day)
median(complete_steps_by_day)
```

## Are there differences in activity patterns between weekdays and weekends?
